---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

DS 202 - lab #5: Drunk Driving and Fatal Accidents

Use the User’s manual to find what the columns stand for. The manual can be found at https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812602

Two of the main tables we are going to look at are the accident and the person tables. These data are made available here: accident, and person

Load the two datasets.
```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(maps)

```

# Question One
Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?

```{r}
p<-acc%>%
  mutate(DAY_WEEK = wday(`DAY_WEEK`,label=TRUE))%>%
  ggplot(aes(x=`DAY_WEEK`))
p+geom_bar()

```

**Accidents are generally highest during the weekend and peak on Saturdays. The fewest accidents occur on Wednesdays.**

# Question 2
Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}
per0<-per%>%
  filter(INJ_SEV==4)
head(per0)
```
# Question 3
Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
per1<-per%>%
  na.omit(MAKE)%>%
  group_by(STATE, MAKE)%>%
  rename(`State Code`= STATE)%>%
  summarize(n=n())%>%
  filter(n==max(n))
  
per1
```
**s

# Question 4
Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map.   (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r}
library(readxl)
states = map_data("state")

textytext<- readxl::read_xlsx("./states.xlsx")%>%
  mutate(`State Name`= tolower(`State Name`)) %>%
  mutate(`State Code`=as.numeric(`State Code`))%>%
  left_join(per1)%>%
  group_by(`State Code`, MAKE)

states1<-states%>%
  group_by(region)%>%
  summarize(avelong=mean(long), avelat=mean(lat), group=median(group))%>%
  left_join(textytext, by=c(region='State Name'))
  

ggplot(states, aes(x='long', y='lat')) + 
  geom_polygon(aes(group = group)) + 
  geom_text(aes(x=avelong, y=avelat, label = MAKE), data = states1)
```

**Although this map is less-than-ideal, the most dangerous vehicle .**

# Question 5
Join the accident and person table (work out which variable(s) to use)

```{r}
baby<-acc %>%
  right_join(per, by='ST_CASE')%>%
  unique()
head(baby)
```
**Handsome baby data frame**

# Question 6
Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.

```{r}
baby0 <- baby %>%
  mutate(SEX = factor(SEX, levels=c(1, 2), labels=c('Male','Female'))) %>%
  na.omit() %>%
  group_by(DAY_WEEK, SEX) %>%
  summarize(`Count of Accidents`=n())

ggplot(baby0, aes(x=DAY_WEEK, y=`Count of Accidents`, fill=SEX, label=`Count of Accidents`)) + geom_bar(stat="identity", position="dodge") + geom_text()
```
**Men are involved in many more accidents than women. Accidents are most frequent for both genders during the weekend.**

```{r}
baby1<- baby%>%
  mutate(SEX = factor(SEX, levels=c(1, 2), labels=c('Male','Female'))) %>%
  na.omit() %>%
  filter(HOUR.x %in% 1:24)%>%
  group_by(HOUR.x, SEX) %>%
  summarize(`Count of Accidents`=n())

ggplot(baby1, aes(x=HOUR.x, y=`Count of Accidents`, fill=SEX, label=`Count of Accidents`)) + geom_bar(stat="identity", position="dodge") + geom_text()
```
**Men crash more than women during all hours of the day, particularly real early in the morning. Accidents peak for both genders around rush hour.**

# Question 7
Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.

```{r}
counties = map_data("county")

acc2<-acc%>%
  select(STATE, FATALS, COUNTY)%>%
  filter(FATALS!=0)%>%
  group_by(STATE, COUNTY)%>%
  summarize(sum=sum(FATALS))



zzz<- textytext%>%
  mutate(subregion = tolower(`County Name`), STATE=as.numeric(`State Code`), COUNTY=as.numeric(`County Code`))%>%
  select(subregion, STATE,`COUNTY`)%>%
  left_join(acc2, by=c('STATE','COUNTY'))%>%
  group_by(STATE, COUNTY)%>%
  unique()%>%
  left_join(counties)%>%
  unique()

ggplot(zzz, aes(x=long, y=lat, fill=sum))+geom_polygon(aes(group=group))

```
**The LA Metro generally has the highest number of per county accidents.**

# Question 8
Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}
#zonked <- textytext%>%
 # mutate(STATE=as.numeric(`State Code`))%>%
  #select(`State Name`, STATE)%>%
  #left_join(acc)%>%
#  filter(FATALS!=0)

#zonked
```



