---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

DS 202 - lab #5: Drunk Driving and Fatal Accidents

Use the User’s manual to find what the columns stand for. The manual can be found at https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/812602

Two of the main tables we are going to look at are the accident and the person tables. These data are made available here: accident, and person

Load the two datasets.
```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(maps)

```

# Question One
Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?

```{r}
p<-acc%>%
  mutate(DAY_WEEK = wday(`DAY_WEEK`,label=TRUE))%>%
  ggplot(aes(x=`DAY_WEEK`))
p+geom_bar()

```

**Accidents are generally highest during the weekend and peak on Saturdays. The fewest accidents occur on Wednesdays.**

# Question 2
Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)

```{r}
per0<-per%>%
  filter(INJ_SEV==4)
head(per0)
```
# Question 3
Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
per1<-per%>%
  na.omit(MAKE)%>%
  group_by(STATE, MAKE)%>%
  summarize(n=n())%>%
  filter(n==max(n))
  
per1
```
# Question 4
Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map.   (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)

```{r}
library(readxl)
textytext <- readxl::read_xlsx(./states.xlsx)
states = map_data("state")
counties = map_data("county")

ggplot(states, aes(x=long, y=lat))+geom_path(aes(group=group))+geom_polygon(aes(group=group))



```
# Question 5
Join the accident and person table (work out which variable(s) to use)

```{r}

```
# Question 6
Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.

```{r}

```
# Question 7
Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.

```{r}

```
# Question 8
Is summer or winter more dangerous? Does this depend on states? Explore and explain.

```{r}

```



